Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction

### 2024-02-25 paper/paper.tex
updated manuscript to address reviewers' comments

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Inland waters pose a unique challenge for water quality monitoring by remote sensing techniques due to their complicated spectral features and small-scale variability. At the same time, collecting the high-quality reference data needed to calibrate remote sensing data products is both time consuming and expensive. In this study, we present the further development of a robotic team composed of an uncrewed surface vessel (USV) providing in situ reference measurements and an unmanned aerial vehicle (UAV) equipped with a hyperspectral imager. Together, this team is able to address the limitations of existing approaches by enabling the simultaneous collection of comprehensive in situ data together with high-resolution hyperspectral imagery. We showcase the capabilities of this team using data collected in a north Texas pond during three collections in 2020. Machine learning models for 13 variables are trained using the data set of paired in situ measurements and coincident reflectance spectra. These models can successfully estimate physical variables, including temperature, conductivity, pH, and turbidity, as well as the concentrations of blue-green algae, colored dissolved organic matter (CDOM), chlorophyll-a, crude oil, optical brighteners and the ions $\mathrm{Ca}^{2+}$, $\mathrm{Cl}^{-}$, and $\mathrm{Na}^{+}$. We extend the training procedure to utilize conformal prediction to estimate 90\% confidence intervals for the output of each trained model. Maps generated by applying each trained model to the collected images reveal the small-scale spatial variability within the pond. Additionally, the permutation importance computed for each feature of the trained models highlights the spectral features relevant to each water quality variable.
}

@@

For decades, remote sensing imagery have been used for environmental monitoring with applications ranging from resource mapping, land type classification, and urban growth assessment to wildfire monitoring, natural disaster tracking, and many more \cite{melesse2007remote, joyce2009review}. Among these applications, the retrieval of water quality variables from remote sensing imagery remains challenging due to the difficulty of obtaining in situ reference data coincident with available satellite imagery. Traditional approaches to obtain these data have relied on serendipitous satellite passes over fixed sensing sites or sensor-equipped vessels. As a consequence, curating comprehensive datasets can require decades of observations \cite{aurin2018remote, ross2019aquasat}. This poses a significant challenge for assessing natural and anthropogenic changes to water composition in real time.

Where datasets have been compiled, many studies have shown success extracting optically active water quality variables such as colored dissolved organic matter (CDOM), chlorophyll-a, and suspended sediment concentration by using combinations of spectral bands from from remote sensing satellites \cite{remote-sensing-finland,bonansea2015using}. These approaches can be further augmented by machine learning methods which consist of nonlinear and non-parametric models designed to learn representations of functions directly from data \cite{sagan2020monitoring}. For example Petersen et al. utilized a deep neural network to successfully estimate blue-green algae, chlorophyal-a, CDOM, as well as the optically in-active

which are often related indirectly to optically active variables.

which  \cite{sagan2020monitoring, peterson2020deep}

can augment traditional approaches using reflectance band ratios to enable estimation of water quality variables and perform water quality classification \cite{schiller1999neural, brown2008neural, lary2010artificial, feng2015urban,belgiu2016random}.

However, the low spatial and spectral resolution of available multiband remote sensing satellites makes it difficult to analyze inland waters with small spatial features and complicated spectral signatures. Furthermore, the long return times of these satellites limit their use for real-time monitoring, for example, during oil spills \cite{fingas2017review}.

Advances in multispectral and hyperspectral imaging technology have considerably reduced their size, making it possible to incorporate these cameras into the payload of autonomous aerial vehicles (UAVs) such as quadcopters, hexacopters, and other drones \cite{hruska2012radiometric}.

Bringing the imaging technology closer to the ground enables UAVs to address the spectral, spatial, and temporal shortcomings of remote sensing satellites.

Controlling altitude during flight, UAVs can produce imagery with pixels at the centimeter scale and limit the need for complicated atmospheric corrections to account for scattering by atmospheric aerosols and gases \cite{adao2017hyperspectral,banerjee2020uav}. Already, UAVs equipped with multispectral and hyperspectral imagers are being used in a variety of domains to great effect, including biomass estimation, forest management, precision agriculture, and now water quality monitoring \cite{adao2017hyperspectral, padua2017uas,arroyo2019implementation,kurihara2020unmanned,ehmann2019monitoring, lu2021retrieval}. Despite the increased spectral and spatial resolution afforded by UAV platforms, these improvements alone cannot address the limited spatial resolution of in situ reference data.

To address this gap and enable comprehensive real-time water quality \textcolor{red}{detection}, we have developed a robotic team composed of an autonomous uncrewed surface vessel (USV) equipped with a collection of reference grade instruments together with a UAV carrying a hyperspectral imager \cite{robotTeam1}. By incorporating these reference instruments on a maneuverable USV platform, we are able to rapidly collect large volumes of water quality data for a variety of physical, biochemical, and chemical variables. Critically, the USV enables the collection of reference data with significantly improved spatial resolution compared to other approaches. These in situ measurements then provide the ground-truth data used to train machine learning models which map reflectance spectra captured by the hyperspectral imager to the desired water quality variables.

In this study, we expand on our previous work by demonstrating the capabilities of this coordinated robot team using data collected at a pond in North Texas over three separate days in 2020. Specifically, we show the ability to accurately predict physical variables such as temperature, conductivity, pH, and turbidity, as well as biochemical constituents, including blue-green algae pigments, CDOM, and chlorophyll-a, in addition to concentrations of crude oils, optical brighteners, and a variety of ions. In addition, the considerable size of the collected data set allows us to quantify the uncertainty in the resulting machine learning models by using conformal prediction techniques. This methodology addresses the shortcomings of existing approaches by facilitating rapid, real-time data collection coupled with on-board hyperspectral image processing and machine learning model deployment with uncertainty quantification. The resulting models are small enough to be deployed on board the UAV to enable real-time streaming of data products to the ground. This approach drastically reduces the time from data collection to insight.

\textcolor{red}{The primary goal of this study is to demonstrate the capabilities of a coordinated robotic team combining in situ sensing together with hyperspectral imaging for the rapid characterization of water composition. Specifically, we seek to address the primary challenge of acquiring remote sensing data products coincident with reference measurements that are critical for the retrieval of water composition and water quality parameters. By bringing together
- there are three key goals in this study: The first is to demonstrate the variety of water quality and composition variables that can be predicted from our collected hypserspectral imager. The second is to demonstrate that hyperspectral imagery from multiple observations can by combined despite variability in lighting conditions, and the third is to introduce a method for uncertainty quantification which is crucial for the evaluation of model predictions. 
}

@@

\textcolor{red}{Should we add something here about non-optically active substances?}
\textcolor{blue}{Good idea if you have something}

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. To model the data, we chose to use the Random Forest Regressor (RFR) as implemented in the Machine Learning Framework for Julia package,  MLJ \cite{MLJ1}.  Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the classification and regression trees (CART) algorithm. Each tree in a RFR is trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}. Random Forests are particularly attractive due to their fast training and inference times. Furthermore, studies continue to observe that tree-based models like the Random Forest remain superior for tabular datasets \cite{grinsztajn2022tree, shwartz2022tabular}.

@@

\begin{figure}[p]
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.2in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-ranking.pdf}
\caption{Ranked permutation importance for each feature in the physical variable models. Permutation importance measured the decrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=\columnwidth]{paper/figures/results/maps/physical.pdf}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

@@

% \cite{cohen2004landsat, melesse2007remote, allison2016airborne, joyce2009review, robert2017analysis}

\textcolor{red}{test
- We need to discuss how we are building on the original paper. 
- The key advancements are a demonstration of more optically active substances that can be successfully modelled from the HSIs 
- The ability to model physical parameters that are not considered to be optically active in the sense that they have uniquelly identifiable absorption features (ions, etc...) 
- The ability to successfully combine mutliple observations
    - This is particularly noteworthy as there was significant variation in total downwelling intensity between the separate collections which could introduce bias between reflectance values measured on separate days
    - Key result is that by incorporating the relevant information (solar illumination geometry and viewing geometry) we are able to successfully account for these biases and produce models that are consistent between multiple observation periods
- Estimation of uncertainty using conformal prediction
- The trained models are small enough so as to be deployable on the UAV for real time processing

- part one: overview of method and description of how we are able to collect comparable data quantities which traditionally require decades i.e. Aurin et al. We can also mention how existing approaches using HSI on UAV platforms suffer from a lack of comprehensive in-situ data needed to reliably calibrate the (machine learning) data products. By creating a coordinated team with both autonomous in situ data collection and rapid HSI acquisition by a UAV we effectively bridge this gap by enabling the precise collection of collocated data points.
- part two: We extend our previous work in three key ways: first we model significantly more target variables, second we confirm that it is possible to utilize imagery from multiple collections despite the impact of light variability so long as we appropriately augment the data to include illumination and viewing geometry, third we implemented a method for robust uncertainty quantification by using conformal prediction so that we can evaluate the certainty of our trained models.
- finally we can comment on limitations, extensions, and future work
    - Limitation: total area surveyed is determined by battery life (suggestion --> multiple drones) 
    - Limitation: predictions of trained tree-based models are litmied by the range of available data. To extend the model to be sensitive across a broader range of values/concentrations, we need to collect more training data. However we have demonstrated that data from disparate observations can be combined despite the limitations of lighting variability. Indeed it is preciesly the fact that we are interested in the water composition and not the absolute reflectance values themselves that makes this possible because so long as we include enough descriptive variables which capture the lighting dynamics, we can account for differences in ambient lighting conditions between observations.
}

Much effort has recently been spent on the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for monitoring water quality. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. have combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other variables of water quality together with Landsat historical reflectance data for the period 1984--2019 \cite{ross2019aquasat}.

In this paper, we have validated the methodology introduced in our previous work\cite{robotTeam1}. Specifically, we demonstrate that by using an autonomous team of sensing sentinels, we can
\begin{enumerate}
    \item Rapidly acquire large volumes in in-situ reference data points for a variety of relevant water quality variables including physical measurements such as temperature, pH, and conductivity as well as ion concentrations, chemicals such as crude oil, and relevant biochemical variables such as blue-green algae pigments, chlorophyll-a, and CDOM. 
    \item Simultaneously collect hyperspectral remote sensing imagery using an UAV.
    \item Rapidly process captured imagery into georectified reflectance data cubes.
    \item Train machine learning models to map captured reflectance spectra to water quality variables measured by reference sensors.
    \item Apply trained models to rapidly map large bodies of water and identify areas of interest.
\end{enumerate}
Additionally, we have demonstrated that we can extend this methodology in two new ways. First, we have shown that it is possible to combine data sets from separate collection periods and successfully train models using \textit{all} of the records despite variability in incident lighting conditions that affects the assumptions made in the reflectance conversion of captured hyperspectral data cubes. Second, we have shown that by leveraging the large amount of data collected, we can simultaneously train machine learning models and quantify their prediction uncertainties by utilizing conformal prediction.

As Figure~\ref{fig:downwelling-hist} indicates, there is variation in the total downwelling intensity between images collected on the same day \textit{and} between each separate collection period despite acquiring the imagery consistently near solar noon. These within-collection variations are due to a combination of the stability of the UAV (on which the upward facing downwelling irradiance spectrometer is mounted) together with the occasional interference of clouds. Moreover, the assumption that the water's surface can be treated as Lambertian is clearly violated when the water is not perfectly still. Despite the potential impact of these limitations on the quality of the resulting reflectance data cubes, the smoothness of the maps generated by our models suggests that we have provided sufficient context by including the relevant solar illumination and viewing geometries as additional features in the final data set.  This fact is reinforced by the position of these variables as the most important features for \textit{each} of the estimated water quality variables. As long as we are primarily interested in these values and not the reflectances themselves, we are able to successfully account for these lighting effects.

With respect to the real-time deployment of these models in the field, we note that training a reduced feature model without further hyperparameter optimization takes roughly 1 minute per target variable of interest using the processing computer included on the UAV. This means that in principle the necessary training data can be collected by the USV, hyperspectral imagery can be acquired and processed into reflectance data cubes by the UAV, the resulting data sets can be combined to identify coincindent records, and this collection can then be used to train machine learning models and generate maps as in Figure~\ref{fig:map-biochem} all while the investigators are still in the field. Analyzing these resulting maps enables areas of interest to be readily pinpointed, as demonstrated by the identification of elevated crude oil, optical brighteners, and CDOM in the eastern alcove of the pond on 11-23. Having identified a region of interest, the USV can then be sent to collect additional reference data and confirm the presence of any spill, bloom, or other source.

Beyond the potential applications of this robot team for direct water quality monitoring, the combination of reference instruments together with easily co-located hyperspectral imagery provides fertile ground for the development of new spectral indices targeted towards specific water quality variables. In this paper, we have shown how permutation importance ranking for trained machine learning models enables a straightforward interpretation of the relative value of each reflectance bin to the final model predictions. In future work, we plan to utilize this information to construct new combinations of spectral bands which can be readily extended for use remote sensing platforms equipped with hyperspectral imagers. \textcolor{red}{The recently launched Environmental Mapping and Analysis Program (EnMAP) is one such example which includes over 91 spectral bands in the VNIR that overlap with our hyperspectral imager \cite{Storch2023TheEI}.}

@@

In this study, we address the two key limitations of current remote sensing approaches to characterizing water quality, namely the limited spatial, spectral, and temporal resolution provided by existing satellite platforms and the lack of comprehensive, highly resolved in-situ data needed to validate remote sensing data products. By equipping an autonomous USV with a suite of reference sensors, we can collect significantly more data than existing approaches that rely on the collection of individual samples for lab analysis or are constrained to continuous sensing at fixed sites. Utilizing an autonomous UAV equipped with a hyperspectral imager in tandem with the USV allows us to rapidly generate aligned datasets which can be used to train machine learning models which map measured reflectance spectra to the desired water quality variables of interest. Furthermore, by virtue of collecting large training datasets, we are able to simultaneously estimate the uncertainty of our trained models using conformal prediction. This addresses the lack of uncertainty quantification in traditional machine learning approaches. Finally, the rapid hyperspectral data cube processing pipeline employed on board the UAV makes it possible to deploy these trained models in order to rapidly generate maps of the desired target variables across entire bodies of water. The rapid turnaround time from data collection to model deployment is critical for real-time water quality assessment and threat identification.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}\label{Model Hyperparameters}
%% \subsection[\appendixname~\thesubsection]{}
\begin{table}[H] 
\caption{Final hyperparameter values for each target model. The number of trees and sampling ratio were optimized using a random search. The maximum tree depth was fixed to 20 to limit overfitting and control the size of the final model. The number of sub-features was set to the square root of the total number of features and the minimum samples per leaf and minimum samples per split were left to their default values. \label{tab:hyperparameters}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCCCC}
\toprule
\textbf{Target}	& \textbf{Number of Trees}	& \textbf{Sampling Ratio} & \textbf{Maximum Tree Depth} & \textbf{Number of Sub-features} & \textbf{Minimum Samples Per Leaf} & \textbf{Minimum Samples Per Split}\\
\midrule
Temperature & 153 & 0.979 & 20 & 5 & 1 & 2 \\
Conductivity & 153 & 0.979 & 20 & 5 & 1 & 2\\
pH & 153 & 0.979 & 20 & 5 & 1 & 2\\
Turbidity & 153 & 0.979 & 20 & 5 & 1 & 2\\
$\mathrm{Ca}^{2+}$ & 153 & 0.979 & 20 & 5 & 1 & 2\\
$\mathrm{Cl^-}$ & 153 & 0.979 & 20 & 5 & 1 & 2\\
$\mathrm{Na^+}$ & 153 & 0.979 & 20 & 5 & 1 & 2\\
%Blue-Green Algae (Phycoerythrin) & 153 & 0.979 & 20 & 5 & 1 & 2\\
Phycoerythrin & 153 & 0.979 & 20 & 5 & 1 & 2\\
CDOM & 153 & 0.979 & 20 & 5 & 1 & 2\\
Chlorophyll A & 153 & 0.979 & 20 & 5 & 1 & 2\\
%Blue-Green Algae (Phycocyanin) & 153 & 0.979 & 20 & 5 & 1 & 2\\
Phycocyanin & 153 & 0.979 & 20 & 5 & 1 & 2\\
Crude Oil & 153 & 0.979 & 20 & 5 & 1 & 2\\
Optical Brighteners & 153 & 0.979 & 20 & 5 & 1 & 2\\
\bottomrule
\end{tabularx}
\end{table}

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Inland waters pose a unique challenge for water quality monitoring by remote sensing techniques due to their complicated spectral features and small-scale variability. At the same time, collecting reference data needed to calibrate remote sensing data products is both time consuming and expensive. In this study, we present the further development of a robotic team composed of an uncrewed surface vessel (USV) providing in situ reference measurements and an unmanned aerial vehicle (UAV) equipped with a hyperspectral imager. Together this team is able to address the limitations of existing approaches by enabling the simultaneous collection of hyperspectral imagery with precisely collocated in situ data. We showcase the capabilities of this team using data collected in a north Texas pond across three days in 2020. Machine learning models for 13 variables are trained using the data set of paired in situ measurements and coincident reflectance spectra. These models successfully estimate physical variables including temperature, conductivity, pH, and turbidity, as well as the concentrations of blue-green algae, colored dissolved organic matter (CDOM), chlorophyll-a, crude oil, optical brighteners and the ions $\mathrm{Ca}^{2+}$, $\mathrm{Cl}^{-}$, and $\mathrm{Na}^{+}$. We extend the training procedure to utilize conformal prediction to estimate 90\% confidence intervals for the output of each trained model. Maps generated by applying the models to the collected images reveal small-scale spatial variability within the pond. This study highlights the value of combining real-time, in situ measurements together with hyperspectral imaging for the rapid characterization of water composition.
}

@@

For decades, remote sensing imagery have been used for environmental monitoring with applications ranging from resource mapping, land type classification, and urban growth assessment to wildfire monitoring, natural disaster tracking, and many more \cite{melesse2007remote, joyce2009review}. Among these applications, the retrieval of water quality variables from remote sensing imagery remains challenging due to the difficulty of obtaining in situ reference data coincident with available satellite imagery. Traditional approaches to obtain these data have relied on serendipitous satellite passes over fixed sensing sites or sensor-equipped vessels. As a consequence, curating comprehensive datasets can require decades of observations \cite{aurin2018remote, ross2019aquasat}. This poses a significant challenge for assessing natural and anthropogenic changes to water composition in real time, for example, during oil spills \cite{fingas2017review}.

Where datasets have been compiled, many studies have shown success extracting optically active water quality variables such as colored dissolved organic matter (CDOM), chlorophyll-a, and suspended sediment concentrations by using combinations of spectral bands from from remote sensing satellites \cite{remote-sensing-finland,bonansea2015using}. These approaches can be further augmented by machine learning methods which consist of nonlinear and non-parametric models designed to learn representations of functions directly from data \cite{lary2010artificial}. For example, Petersen et al. utilized a deep neural network to successfully estimate blue-green algae, chlorophyl-a, CDOM, dissolved oxygen, specific conductance, and turbidity from a fused dataset of Landsat-8 and Sentinel-2 imagery \cite{peterson2020deep}. Other methods such as support vector machines and decision trees have also been successfully applied to retrieve water quality parameters from remote sensing imagery \cite{belgiu2016random,sagan2020monitoring}. However, the low spatial and spectral resolution of available multiband remote sensing satellites makes it difficult to analyze inland waters with small spatial features and complicated spectral signatures.

Advances in multispectral and hyperspectral imaging technology have led to a considerable reduction in size making it possible to incorporate these cameras into the payload of autonomous aerial vehicles (UAVs) such as quadcopters, hexacopters, and other drones \cite{hruska2012radiometric}. Flying at low altitudes enables the collection of centimeter scale imagery whilst limiting the need for complicated atmospheric corrections to account for scattering by atmospheric aerosols and gasses \cite{adao2017hyperspectral, banerjee2020uav}. Already UAVs equipped with multispectral and hyperspectral imagers are being used in a variety of domains to great effect including biomass estimation, forest management, precision agriculture, and recently, water quality monitoring \cite{adao2017hyperspectral, padua2017uas,arroyo2019implementation,kurihara2020unmanned, ehmann2019monitoring}. Despite the superior spectral and spatial resolution enabled by UAV platforms, these improvements alone do not address the limited spatial coverage of in situ reference data used for associated water composition retrieval. For instance, Lu et al. used a UAV-born hyperspectral imager to develop machine learning models for the inversion of chlorophyl-a and suspended solids using samples from 33 fixed locations \cite{lu2021retrieval}. Similarly, Zhang et al utilized a UAV equipped with a hyperspectral imager to estimate water quality parameters by collecting imagery coincident with samples taken from 18 sampling sites \cite{zhang2022selection}. In both of these examples, the collection of in situ reference data remains the key challenge for the application of UAV systems to water quality quantification.

To address this gap and enable comprehensive, real-time evaluation of water composition, we have developed a robotic team comprised of an autonomous uncrewed surface vessel (USV) equipped with a collection of reference grade instruments together with a UAV carrying a hyperspectral imager. By incorporating reference instruments on a maneuverable USV platform, we are able to rapidly collect large volumes of water quality data for a comprehensive suite of physical, biochemical, and chemical variables precisely collocated with spectra captured by the UAV. Critically, the USV enables the collection of reference data with significantly improved spatial resolution compared to other approaches. In our previous work, we introduced this paradigm and described how in situ measurements collected by the USV can be used to provide the ground-truth data for machine learning models which map reflectance spectra captured by the hyperspectral imager to the desired water quality variables \cite{robotTeam1}.

The main objective of this study is to expand on our previous work in three new ways: The first is to explore the breadth of water quality variables that can be inferred from collected hyperspectral imagery. With the goal of comprehensive measurement in mind, we demonstrate the ability to accurately predict physical variables such as temperature, conductivity, pH, and turbidity, as well as biochemical constituents including blue-green algae pigments, CDOM, and chlorophyll-a, in addition to concentrations of crude oils, optical brighteners, and a variety of ions. The second addition is to demonstrate that observations from separate collections can effectively be combined by carefully accounting for variability in the viewing and illumination geometries of the scene. Finally, we expand our machine learning approach to take advantage of the considerable volume of collected data in order to determine reliable confidence intervals for each predicted parameter using conformal prediction.

@@

The inclusion of optically inactive variables such as conductivity, pH, and ion concentrations was motivated by a desire to be comprehensive. Multiple studies have classified inland water bodies according to their ionic compositions \cite{piper1944graphic, dordoni2023preliminary}. Other research indicates that the structure of dissolved organic matter is affected by changes in pH and cation concentration\cite{pace2012ph}. Therefore changes in physical parameters and ionic content can be expected to be related to the observed distribution of optically active variables in the water. It is therefore reasonable to expect that these parameters may be estimated from hyperspectral images at a given site.

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. To model the data, we chose to use the Random Forest Regressor (RFR) as implemented in the Machine Learning Framework for Julia package,  MLJ \cite{ben_sadeghi_2022_7359268, MLJ1}.  Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the classification and regression trees (CART) algorithm. Each tree in a RFR is trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}. Random Forests are particularly attractive due to their fast training and inference times. Furthermore, studies continue to observe that tree-based models like the Random Forest remain superior for tabular datasets \cite{grinsztajn2022tree, shwartz2022tabular}.

@@

@@

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

\begin{figure}
\centering
\vspace{-0.2in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-ranking.pdf}
\caption{Ranked permutation importance for each feature in the physical variable models. Permutation importance measured the decrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

\begin{figure}
\centering
\vspace{-0.3in}
\includegraphics[width=\columnwidth]{paper/figures/results/maps/physical.pdf}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

\newpage

@@

In recent years, much effort has been spent on the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for retrieving water quality parameters. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other water quality variables with historical Landsat reflectance data for the period 1984--2019 \cite{ross2019aquasat}. The sensing paradigm we have demonstrated here was able to rapidly collect comparable volumes of data within the span of just three observation periods. Therefore, despite the fact that individual UAV tracks cover far less spatial extent than remote sensing imagery, the ability to collect coordinated in situ measurements together with detailed hyperspectral images offers a significant improvement over these traditional approaches. With a coordinated robot team, one does not need to rely on infrequent satellite overpasses when planning data collection. Further, the time offset between reference measurements and remote sensing is significantly reduced from days to minutes.

This study is not the first to employ UAVs equipped with multispectral or hyperspectral imagers for the purpose of assessing water composition and quality. Indeed there are many such examples focused on assessing optically active and inactive water quality parameters using band ratios and machine learning methods \cite{vogt2016near,lu2021retrieval, zhang2022selection}. The key advancement demonstrated by our robot team is the ability to combine UAV-borne hyperspectral imagers together with comprehensive, autonomous in situ sensing for a significant improvement in data volume. Removing the need to collect individual samples, often by hand, for the calibration of water quality models dramatically reduces the time required to perform these analyses. Furthermore, the USV enables rapid validation of model predictions. If a region of water imaged by the UAV predicts elevated levels of a particular water quality parameter, the USV can be provisioned to validate these estimates with its reference sensors.

In \cite{robotTeam1}, we introduced this paradigm. In this study, we have built on this approach in three new ways. First, we have demonstrated the ability to effectively combine observations from disparate collections by augmenting the machine learning models with sufficient features describing the illumination and viewing geometries. As Figure~\ref{fig:downwelling-hist} indicates, we observed variation in the total downwelling intensity between images collected on the same day \textit{and} between each separate collection period. These within-collection variations are due to a combination of the stability of the UAV (on which the upward facing downwelling irradiance spectrometer is mounted) together with the occasional interference of clouds. Moreover, the assumption that the water's surface can be treated as Lambertian is clearly violated when the water is not perfectly still. Despite the potential impact of these limitations on the quality of the resulting reflectance data cubes, the smoothness of the maps generated by our models suggests that we have provided sufficient context by including the relevant solar illumination and viewing angles as additional features in the final data set.  This fact is reinforced by the position of these variables as the most important features for \textit{each} of the estimated water quality variables. As long as we are primarily interested in these values and not the reflectances themselves, we are able to successfully account for these lighting effects when combining data from multiple collections.

The second contribution of this study is to explore the breadth of possible water quality and composition parameters that can be accurately mapped by hyperspectral imagery collected by the UAV. The results presented here confirm the ability of the robot team to predict optically active parameters including blue-green algae, chlorophyll-a, CDOM, crude oil, optical brighteners, turbidity, and temperature. Additionally, we are also also able to infer the distributions of optically inactive variables including conductivity, pH, and ion concentrations. Other studies using multispectral and hyperspectral remote sensing imagery have also estimated optically inactive water quality parameters with the ability to do so stemming from the relationship of these variables to optically active properties of the water \cite{vakili2020determination,guo2021machine,niu2021deep}. We note that in our investigation, the models trained for many of these variables often outperformed their optically active counterparts. As the abundance of these variables is likely tied to the specific composition and content of the pond, it is unlikely that models trained for these optically inactive variables can generalized to other bodies of water.

The third contribution of this work is the extension of our machine learning approach to include uncertainty quantification via conformal prediction. For water quality risk assessment, the trustworthiness of model predictions is of equal or greater importance to the values themselves. However, robust uncertainty quantification has historically been challenging for many machine learning models which behave like black boxes. By leveraging the large data volume collected by the robot team, we are able to simultaneously train predictive models and evaluate confidence intervals for their predictions. As the final column of Table~\ref{tab:fit-results} confirms, the empirical coverage on the holdout testing set provided by the inferred confidence intervals achieves the desired coverage to within 1\%. We chose to use a 90\% confidence interval for this study, but this can easily be adapted to suite the needs of a specific application if greater confidence is required.

With respect to the real-time deployment of these models in the field, we note that training a reduced feature model without further hyperparameter optimization takes roughly 1 minute per target variable of interest using the processing computer included on the UAV. This means that, in principle, training data can be collected by the USV, imagery can be acquired and processed by the UAV, the resulting data sets can be combined to identify coincident records, and this collection can then be used to train machine learning models and generate maps all while the investigators are still in the field. Analyzing these resulting maps enables areas of interest to be readily pinpointed, as demonstrated by the identification of slightly elevated crude oil, optical brighteners, and CDOM in the eastern alcove of the pond on 11-23.

Finally, we note that the high spectral resolution of the UAV imagery together with the ability to collect precisely co-located reference measurements provides a fertile ground for the development of new spectral indices targeted towards water quality variables. In this paper, we have shown that permutation importance ranking for trained machine learning models enables a straightforward interpretation of the relative value of each reflectance bin to the final model predictions. In future work, we plan to utilize this information to identify combinations of spectral bands which can be applied to remote sensing imagery captured by satellites equipped with hyperspectral imagers. The recently launched Environmental Mapping and Analysis Program (EnMAP) is one such example which includes over 91 spectral bands in the VNIR that overlap with those of our hyperspectral imager \cite{Storch2023TheEI}.

@@

In this study, we address two key limitations of current remote sensing approaches to characterizing water quality, namely the limited spatial, spectral, and temporal resolution provided by existing satellite platforms and the lack of comprehensive in-situ measurements needed to validate remote sensing data products. By equipping an autonomous USV with a suite of reference sensors, we rapidly collect significantly more data than existing approaches which rely on the collection of individual samples for lab analysis or are constrained to continuous sensing at fixed sites. Utilizing an autonomous UAV equipped with a hyperspectral imager in tandem with the USV allows us to quickly generate aligned datasets which are used to train machine learning models mapping measured reflectance spectra to the desired water quality variables. By virtue of this increased data volume, we are able to simultaneously estimate the uncertainty of our models by using conformal prediction. Finally, the hyperspectral data cube processing pipeline employed on board the UAV makes it possible to deploy these trained models to swiftly generate maps of the target variables across bodies of water. The rapid turnaround time from data collection to model deployment is critical for real-time water quality evaluation and risk assessment.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}\label{Model Hyperparameters}
%% \subsection[\appendixname~\thesubsection]{}
\begin{table}[H] 
\caption{Final hyperparameter values for each target model. The number of trees and sampling ratio were optimized using a random search. The maximum tree depth was fixed to 20 to limit overfitting and control the size of the final model. The number of sub-features was set to the square root of the total number of features and the minimum samples per leaf and minimum samples per split were left to their default values. \label{tab:hyperparameters}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCCCC}
\toprule
\textbf{Target}	& \textbf{Number of Trees}	& \textbf{Sampling Ratio} & \textbf{Maximum Tree Depth} & \textbf{Number of Sub-features} & \textbf{Minimum Samples Per Leaf} & \textbf{Minimum Samples Per Split}\\
\midrule
Temperature & 153 & 0.979 & 20 & 5 & 1 & 2 \\
Conductivity & 154 & 0.992 & 20 & 5 & 1 & 2\\
pH & 103 & 0.972 & 20 & 5 & 1 & 2\\
Turbidity & 158 & 0.998 & 20 & 5 & 1 & 2\\
$\mathrm{Ca}^{2+}$ & 172 & 0.984 & 20 & 5 & 1 & 2\\
$\mathrm{Cl^-}$ & 110 & 0.999 & 20 & 5 & 1 & 2\\
$\mathrm{Na^+}$ & 103 & 0.972 & 20 & 5 & 1 & 2\\
%Blue-Green Algae (Phycoerythrin) & 153 & 0.979 & 20 & 5 & 1 & 2\\
Phycoerythrin & 158 & 0.998 & 20 & 5 & 1 & 2\\
CDOM & 157 & 0.982 & 20 & 5 & 1 & 2\\
Chlorophyll A & 158 & 0.998 & 20 & 5 & 1 & 2\\
%Blue-Green Algae (Phycocyanin) & 153 & 0.979 & 20 & 5 & 1 & 2\\
Phycocyanin & 142 & 0.995 & 20 & 5 & 1 & 2\\
Crude Oil & 154 & 0.992 & 20 & 5 & 1 & 2\\
Optical Brighteners & 157 & 0.982 & 20 & 5 & 1 & 2\\
\bottomrule
\end{tabularx}
\end{table}
